%!TEX encoding = UTF-8

\chapter*{Acknowledgements}

% get inspiration from
% http://www.robots.ox.ac.uk/~teo/thesis/Thesis/deCampos3DHandTracking1.pdf

First of all, I would like to thank my daily supervisor in this project for his uninterrupted support and patience during my eight months of stay in VisLab and Institute for Systems and Robotics, Instituto Superior T\'{e}cnico, Lisbon. Thank you very much, Prof.~Alexandre~Bernardino. That, plus\dots{} passionately discussing algorithms while eating a Portuguese \emph{doce} and sipping a coffee together (several times) is priceless.

I also wish to express my gratitude to my home advisor: Prof.~Daniele~Nardi of Sapienza University of Rome. Not only has he provided me with the chance to do my thesis research abroad, but he has been helpful, encouraging and available for suggestions at all times.

I would like to thank Prof.~Jos\'{e}~Santos-Victor of Instituto Superior T\'{e}cnico for his confidence in hosting me, and for making VisLab such an enjoyable environment to work at, ultimately making it a breeze to do research there. %

This work was partially supported by EC Project IST-004370 RobotCub and by the Portuguese Government~--~\begin{otherlanguage*}{portuguese}Funda\c{c}\~{a}o para a Ci\^{e}ncia e Tecnologia\end{otherlanguage*} (ISR/IST pluriannual funding) through the POS\_Conhecimento Program that includes FEDER funds. This is gratefully acknowledged.

\bigskip

In Lisbon I found plenty of nice people since day one. I am glad to have met the \emph{Italian gang} of IST, dearest friends and valued teachers to me: Alessio, Giampiero and Matteo. And then I certainly wish to thank many more colleagues for the fruitful discussions that ensued and for the fun (apologies if I forgot anybody): Christian, Daniel, Daniela, Dario, Prof.~Gaspar, Ivana, Jonas the Superschwiizer, Jonas the Swede, Luis\~{a}o, L.~Vargas, Manuel, Marco, M\'ario, Matthijs, Plinio, Ricardo, Rub\'{e}n, Samuel, Verica.

Thank you Aleka, Andrea, Patty, Sabrina, Valentin and everybody else back from my Erasmus year. Cheers to Cimpe and Claudione, crazy Portugal-lovers. Oh, thanks to PierFrok for the tip.
% o PierPal?

% Vicky, Lisa, Tomas? Toni, SŽrgio

\begin{otherlanguage*}{portuguese}
Nico (e Tigas), obrigado por tudo, desde 2005\copyright.
Obrigado \`{a} \emph{trindade}: Carmen, Joana, Lara, Leonor. Agrade\c{c}o tamb\'{e}m ao Gustavo, ao Pi\c{c}arra, ao Paulo e a todos os portugueses que andam por Roma.
\end{otherlanguage*}

\bigskip

\begin{otherlanguage*}{italian}
Naturalmente grazie a tutti coloro con cui ho condiviso esperienze in Italia. A Balerio per l'amicizia antica. A Sofia, \emph{sister of a lifetime}. Ai compagni di scuola: Ilardo, Il Giulio, Jessica, Lorenzo, Nausicaa, Valerione Picchiassi. Ai compagni di universit\`{a}: la \emph{crew} ``Gli anni '12'' e lo staff di foruming.

% Flo e famiglia?

Grazie ad Aurora, a Ludovico e ai gatti tutti: anche se preferisco i cani, faccio un'eccezione.

Grazie a tutti gli altri amici relativamente recenti ma gi\`{a} ottimi e abbondanti: Aldo Rock, Mauro, Simone, Riccardo.

\bigskip

E poi, grazie davvero a \emph{tutta} la mia famiglia. A mia madre Paola e a mio padre Francesco per il loro amore infinito. A mio fratello Davide che \`{e} sempre stato fonte di ispirazione; ad Arianna e a Carlo.
\end{otherlanguage*}